---
title: "Course Project"
author: "Fatemeh Abyarjoo"
date: "Friday, June 12, 2015"
output: pdf_document
---

The goal of your project is to predict the manner in which they did the exercise. This is the "classe" variable in the training set. You may use any of the other variables to predict with. You should create a report describing how you built your model, how you used cross validation, what you think the expected out of sample error is, and why you made the choices you did. You will also use your prediction model to predict 20 different test cases. 

LOADING THE DATA
```{r}
library(lattice)
library(ggplot2)
library(caret)
trainurl="http://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv"
testurl="http://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv"

traindata= read.csv(url(trainurl), na.strings=c("NA","#DIV/0!",""))
testdata=read.csv(url(testurl), na.strings=c("NA","#DIV/0!",""))

dim(traindata)
dim(testdata)

#names(traindata)
#head(traindata)
#head(testdata)
```
CLEAN THE DATA

```{r}
# eliminating near zeros
nearz= nearZeroVar(traindata)
traindata= traindata[, -nearz]

nearz= nearZeroVar(testdata)
testdata= testdata[, -nearz]
# eliminating irrelevant columns
traindata=traindata[,-c(1:7)]
testdata=testdata[,-c(1:7)]

# eliminating columns which are all NAs
traindata= traindata[,colSums(is.na(traindata)) == 0]
testdata= testdata[,colSums(is.na(testdata)) == 0]

dim(traindata)
```

PARTITIONING THE TRAINING DATA

```{r}
set.seed(2000)
data1= createDataPartition(y=traindata$classe, p=0.75, list=FALSE)
train_sub= traindata[data1, ] 
test_sub= traindata[-data1, ]

#plot(train_sub$classe, col="green",xlab="classe", ylab="Frequency")
```
MODEL BUILDING:
First model:Decision tree

```{r}
#install.packages("randomForest")
#install.packages("rpart")
#install.packages("rpart.plot")
#install.packages("rattle")
library(randomForest)
library(rpart) 
library(rpart.plot)
library(rattle)
firstmodel= rpart(classe ~ ., data=train_sub, method="class")
fancyRpartPlot(firstmodel)
firstpredict=predict(firstmodel, test_sub, type = "class")

dtree= confusionMatrix(firstpredict, test_sub$classe)
dtree
```

Second model: Random forest
```{r}
secondmodel=randomForest(classe ~ ., data=train_sub)
secondpredict= predict(secondmodel, test_sub, type = "class")

rforest= confusionMatrix(secondpredict, test_sub$classe)
rforest
```







